---
title: "Greenhouse Gas Emissions in the European Union (2019)"
classoption: landscape, twocolumn
geometry: "left=1cm,right=1cm,top=0cm,bottom=1.5cm"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*A. Name of the applicant: Andrea Colombo*

*B-C. Charts*

```{r pressure, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6 , fig.height = 5.5}
# Library Loading
library(dplyr)
library(ggplot2)
library(ggrepel)
library(plotly)
library(stringr)
library(tidyr)
library(ggforce)


# Data Loading

greenhouse <- read.csv("Data/greenhouse.csv")[,-1]
population <- read.csv("Data/population.csv")[,-1]
area <- read.csv('Data/area_km.csv', stringsAsFactors=FALSE) %>% filter(LANDUSE == 'Total area') %>% select(TIME, GEO, Value) %>% distinct() %>% filter(TIME == '2014')

area$Value <- as.numeric(gsub(",.*$", "", area$Value))
area$GEO[5] <- "Germany"
area$Value <- area$Value *1000

names(area)[2] <- "countries"
names(area)[3] <- "area"

# Clean Date
greenhouse$time <- substr(greenhouse$time,1,4)
population$time <- substr(population$time,1,4)

 # Analysis of levels
# levels(greenhouse$unit)
# levels(greenhouse$airpol)
# levels(greenhouse$geo)

 
# Uniform unit of measure

greenhouse[which(greenhouse$unit == 'THS_T'), 'values'] <- greenhouse[which(greenhouse$unit == 'THS_T'), 'values']/1000 

#Drop useless columns and distinct

greenhouse <- greenhouse %>% filter(unit == 'MIO_T') %>% select(-unit) %>% distinct()

# Keep only CO2 equivalents
greenhouse <- greenhouse %>% filter(airpol != 'CH4') %>% filter(airpol != 'N2O')


 # Task 1: population and gas emission

x <- colnames(greenhouse)
x[4] <- "gas"
colnames(greenhouse) <- x

x <- colnames(population)
x[3] <- "population"
colnames(population) <- x

rm(x)

# Filter out non-EU countries 
greenhouse <- greenhouse %>% filter(substring(geo,1,2) != 'EU') %>% filter(substring(geo,1,2) != 'IS')
greenhouse$geo <-as.character(greenhouse$geo)

population$geo <-as.character(population$geo)

green_popu <- greenhouse %>% inner_join(population, by = c('geo', 'time'))
green_popu <- green_popu %>% mutate_if(is.character, as.factor)

green_popu <- green_popu %>% filter(airpol == 'GHG') %>% select(-airpol) 

countries <- c('Austria', 'Belgium', 'Bulgaria', 'Cyprus', 'Czechia', 'Germany', 'Denmark', 'Estonia','Greece', 'Spain', 'Finland', 'France', 'Croatia',
               'Hungary', 'Ireland', 'Italy',  'Lithuania', 'Luxembourg', 'Latvia', 'Malta', 'Netherlands', 'Poland', 'Portugal', 'Romania',  'Sweden', 'Slovenia','Slovakia', 'United Kingdom' )

green_popu$population <- green_popu$population/1000000

green_popu <- data.frame(green_popu, countries, gas_per_pop = green_popu$gas/green_popu$population)


years_available <- green_popu %>% select(time) %>% distinct()

df<- green_popu %>% filter(time == '2019')



green_popu <- green_popu %>% inner_join(area, by = c('countries')) %>% select(-TIME)
green_popu <- data.frame(green_popu, gas_per_area = green_popu$gas/green_popu$area)

dataset <- green_popu %>% filter(time == '2019')

dataset <- data.frame(dataset,zoom = ifelse(dataset$population < 15, TRUE, FALSE))


ggp <- ggplot(dataset, mapping = aes(x=population, y=gas)) +

  geom_text_repel(
   mapping = aes(label = countries),
    size = 2,
    min.segment.length = 0,
    seed = 42,
    box.padding = 0.4,
    arrow = arrow(length = unit(0.007, "npc")),
    nudge_x = .5,
    nudge_y = .5,
    color = "grey60") +
  
    geom_point(data = dataset,aes(colour=countries, size = gas/population), alpha=0.6) +
  
  labs(
    title = 'Greenhouse Gas Emissions and Population (EU countries)',
    subtitle = 'Total Greenhouse gas emissions in CO2 equivalent for the year 2019 (excluding LULUCF and memory items, including international aviation) and population',
    x = "Population (in Millions)",
    y = "Total Greenhouse Emissions (Mil. tonnes)"
  ) +  
  theme(
    legend.position = "none",
    # Customize title and subtitle font/size/color
    plot.title = element_text(
      size = 10, hjust = 0.5,
      color = "#2a475e"
    ),
    
    # Adjust axis parameters such as size and color.
    axis.text = element_text(size = 8, color = "black"),
    axis.title = element_text(size = 7),
    axis.line = element_line(colour = "grey50", arrow = grid::arrow(length = unit(0.3, "cm"))),
    
    # Use a light color for the background of the plot and the panel.
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white"),
    plot.subtitle=element_text(size=5, hjust=0.5, face="italic", color="black")
  )+
  facet_zoom(x = population < 15, y = gas < 190, horizontal = FALSE ,zoom.size = 0.3, show.area = FALSE,
             zoom.data = dataset$zoom)  + coord_cartesian(clip="off")

ggp

```

The size of the points depends on the relative amount of greenhouse gas emissions, compared to population:

$$
Size\; Points = \frac{Total\; greenhouse\; gas\; emissions}{Population}
$$

The same is valid for the second chart, where Population is replaced with the Area


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6 , fig.height = 4}
ggp <- ggplot(dataset, mapping = aes(x=area, y=gas)) +
  geom_point(aes(colour=countries, size = gas/area), alpha=0.6) + 
  geom_text_repel(
    aes(label = countries),
    size = 2,
    min.segment.length = 0,
    seed = 42,
    box.padding = 0.4,
    arrow = arrow(length = unit(0.007, "npc")),
    nudge_x = .5,
    nudge_y = .5,
    color = "grey60") +
  
labs(
  title = 'Greenhouse Gas Emissions and Area (EU countries)',
  x = "Area (KM2)",
  y = "Total Greenhouse Emissions (Mil. of tonnes)",
  subtitle = 'Total Greenhouse gas emissions in CO2 equivalent for the year 2019 (excluding LULUCF and memory items, including international aviation) and area (km2)',
) +  
  theme(
    legend.position = "none",
    # Customize title and subtitle font/size/color
    plot.title = element_text(
      size = 10,hjust = 0.5,
      color = "#2a475e"
    ),

    
    # Adjust axis parameters such as size and color.
    axis.text = element_text(size = 8, color = "black"),
    axis.title = element_text(size = 7),
    axis.line = element_line(colour = "grey50", arrow = grid::arrow(length = unit(0.3, "cm"))),
    
    # Use a light color for the background of the plot and the panel.
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white"),
    plot.subtitle=element_text(size=5, hjust=0.5, face="italic", color="black")
  )+ scale_x_continuous(labels = scales::comma)

ggp
```

*D. Brief analysis*

An interactive version of these analysis, including further ones, has been developed in a R Shiny app, available at https://andreacolombo.shinyapps.io/gasemissionapp/

\begin{enumerate}

\item Direct relationship between tonnes of greenhouse gas emissions and number of people living in a country: the higher is the population the more emissions that country will emit. Looking at the relative terms of this relationship (the size of the points) we can see that some smaller countries are the worst placed, as Luxembourg, which is the worst ranked country with around 20 tonnes of greenhouse gas emissions per millions of people, while the EU mean is 8.28 tonnes.

\item Unclear relationship for chart 2, although an underlying direct relationship may be observed. In relative terms, Finland, Sweden and Malta are the EU countries that emit the less emissions compared to the size of their territories. The worst placed EU countries are the ones with an high population density and host of many industries.

\item A comprehensive analysis of the charts and the underlying data shows that Scandinavian countries are the less polluting EU countries in terms of CO2, with Finland and Sweden performing well in both the charts showed. 


\end{enumerate}